Earthquakes: consequences of disasters
====================================================================

```{r}
setwd("D:/R/Udacity/EDA_Course_Materials/lesson4")
library(ggplot2)
library(tidyr)
library(dplyr)
```

## Affected and killed by earthquakes
```{r}
#read two files
affected <- read.csv("indicator_earthquake affected - Data.csv", header = TRUE, row.names = 1)
killed <- read.csv("indicator_earthquake killed - Data.csv", header = TRUE, row.names = 1)
```

```{r}
#add new columns to both dataset which represent casualties and countries
affected['casualty'] = 'affected'
affected['country'] <- row.names(affected)
killed['casualty'] = 'killed'
killed['country'] <- row.names(killed)
```

```{r}
#making tidydata 
tidy_affected <- gather(affected, year, total, -casualty, - country)
tidy_killed <- gather(killed, year, total, -casualty, - country)
```

```{r}
#merge two datasets
data <- rbind(tidy_affected, tidy_killed)

#correct year names
data['year'] <- extract_numeric(data$year)

#adjust total column by converting into thousands and replacing 0 value with 1
data['total'] <- (data$total/1000) + 1
```

## Explore total number of affected and killed people (in thousands)
We can see that max value of 'total' is very far from mean value. when we make a plot we can see that we have two outliers.
```{r}
summary(data$total)

ggplot(aes(x=year, y = total), data = data) +  
    geom_point() + 
    scale_y_continuous(breaks = seq(0,47180,10000), lim = c(1.002,47180)) +
    scale_x_continuous(breaks=seq(1970,2008,2))
```


## Transform data 
Let's transform data to make it dispersed. Now we can see better distribution of earthquakes in the world
```{r}
ggplot(aes(x=year, y = total), data = data) +
    geom_point() + 
    scale_y_continuous(breaks = seq(0,47180,10000), lim = c(1.002,47180)) +
    scale_x_continuous(breaks=seq(1970,2008,2)) +  
    coord_trans(y = "log10")
```

Let's check correlation between earthquakes and years.

```{r}
cor.test(data$total, data$year)
```
There is no coorelation between earthquakes and years as is was expected.

## Earthquakes by casualty 
When data is grouped by casualty we can see that number of people affected by earthquake is much higher than number of killed people

```{r}
ggplot(aes(x=year, y = total), data = data) +
    geom_point(aes(colour=casualty, group= casualty)) + 
    scale_y_continuous(breaks = seq(0,47180,10000), lim = c(1.002,47180)) +
    scale_x_continuous(breaks=seq(1970,2008,5)) + 
    coord_trans(y = "log10") +
    labs(title = "Earthquakes by casualty")
```

## Top 10 earthquakes by number of affected people (in thousands)
```{r}
top_af <- head(data[order(-data$total),],10)
top_af
```

## Top 10 earthquakes by number of killed people (in thousands)
```{r}
top_killed<-filter(data, casualty == "killed")
top_killed<-head(arrange(top_killed, -total), 10)
top_killed
```

## Most seismoactive countries
Let's create a vector with unique values which represents most seismoactive countries according two top_affected and top_killed
```{r}
danger <- unique(c(top_af$country, top_killed$country))
danger
length(danger)
danger_set<-filter(data, country == danger)
```

Let's plot data of most dangerous seismoactive countries
```{r}
ggplot(aes(x=year, y = total), data = danger_set) +
    geom_point(aes(colour=country, group= country, size = sqrt(total))) +
    scale_size_area() +
    coord_trans(y = "log10") +
    scale_y_continuous(lim = c(1.002,47180)) +
    scale_x_continuous(breaks=seq(1970,2008,5))    
```

We cannot see very well earthquakes' sizes of most countries so let's zoom in our plot and look at affected and killed seperately
```{r}
ggplot(aes(x=year, y = total), data = top_af) +
    geom_point(aes(colour=country, group= country, size = sqrt(total))) +
    scale_size_area() +
    coord_trans(y = "log10") +
    scale_x_continuous(breaks=seq(1970,2008,5)) +
    labs(title = "Earthquakes' size and country, by affected people in thousands")

ggplot(aes(x=year, y = total), data = top_killed) +
    geom_point(aes(colour=country, group= country, size = sqrt(total))) +
    scale_size_area() +
    coord_trans(y = "log10") +
    scale_x_continuous(breaks=seq(1970,2008,5))+
    labs(title = "Earthquakes' size and country, by killed people in thousands")
```


```{r, fig.width = 10, fig.height = 5}
country_set<-filter(data, country == danger)

ggplot(aes(x=year, y = total), data = country_set) +
    geom_point(color = "red") +
    geom_smooth() +
    facet_wrap(~country, ncol=4, scale = "free_y")
```

