
```{r}
setwd("D:/R/Udacity/EDA_Course_Materials/lesson3")
library(rmarkdown)
library(ggplot2)
library(dplyr)
```

# Investigation of my friends' distribution birth months and days

###My aim in this project is to answer the folowing questions:

* Which month contains the most number of birthdays?

* How many birthdays are in each month?

* Which day of the year has the most number of birthdays?

* Do I have at least 365 friends that have birthdays on everyday of the year?

```{r}
csvdata <- read.csv('FBbd.csv')

#convert vector into necessary time format
bdates <- strptime(csvdata$Start, "%d/%m/%Y %H:%M")
names(unclass(bdates))

#extract days and months
days <- bdates$mday
mon <- as.numeric(bdates$mon) + 1
yday <- as.numeric(bdates$yday) + 1

#creat dataframes
df <- data.frame(days, mon, yday, count = 1)
```

##Which month contains the most number of birthdays?
Histogram shows that mostly my friends have birthdays in the range from March to July and most of them were born in March.

```{r}
qplot(df$mon, binwidth = 1,
      main = 'Birthdays per month',
      xlab = 'Months',
      ylab = 'Number of birthdays in month',
      color = I('black'), fill = I('#A41470')) +
    scale_x_continuous(breaks = seq(1,12,1), lim = c(1,13)) +
    scale_y_continuous(breaks = seq(0,30,2), lim = c(0,30))
ggsave('BD_num.png')
```

##How many birthdays are in each month?
```{r}
by_mon<-group_by(df, mon)
summarise(by_mon, sum(count))
```

## Which day of the year has the most number of birthdays?
Among my friends the most number of birthdays is on 1 of May.
```{r}
# convert day of year into "day-month format
by_yday<-group_by(df, yday)
df_yday<-summarise(by_yday, total = sum(count))
df_yday$yday <- format(strptime(df_yday$yday, format="%j"), format="%d-%m")
ordered <- df_yday[order(-df_yday$total),]
```
## Do I have at least 365 friends that have birthdays on everyday of the year?
I sorted days of birth of my friends. Final table shows that I have 252 observations so it already means that my friends' birthdays are not every day of the year.
```{r}
bd <- select(df, days, mon)
bd_ordered <- bd[order(as.numeric(bd$mon), as.numeric(bd$days)),]
row.names(bd_ordered) <- c(1:252)
bd_ordered
```

```{r}
qplot(data = by_yday, x = yday, binwidth = 1,
      main = 'Birthdays during one year',
      xlab = 'Days of year',
      ylab = 'Number of birthday in that day',
      fill = I('#A41470')) +
      scale_x_continuous(breaks = seq(0,365,30), lim = c(1,366))
```

