
```{r}
setwd("D:/R/Udacity/EDA_Course_Materials/lesson3")
```


Lesson 3
============================
## Reading in Data
    
    
```{r}
getwd()
list.files()
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
library(ggplot2)
library(ggthemes)
names(pf)
#theme_set(theme_minimal(24))
#qplot(x = dob_day, data = pf)
ggplot(aes(x = dob_day), data = pf) + 
    geom_histogram() + 
    scale_x_discrete(breaks = 1:31) +
    facet_wrap(~dob_month, ncol=3)

```

## Count friends

```{r}
qplot(x = friend_count, data = pf, xlim = c(0,1000))
# another way to do a histogram:
qplot(x = friend_count, data = na.omit(pf), binwidth = 10) + 
    scale_x_continuous(lim = c(0,1000), breaks = seq(0, 1000, 50)) +
    facet_wrap(~gender)

```


## Statistics 'by' Gender

```{r}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```


## Tenure
```{r}
qplot(x = tenure/365, data = pf, binwidth = .25,
      xlab = 'Number of years using Facebook',
      ylab = 'Number of users in sample',
      color = I('black'), fill = I('#F79420')) +
    scale_x_continuous(lim = c(0,7), breaks = seq(0, 7, 1))
```


## Age
```{r}
qplot(x=age, data = pf, binwidth = 5,
      xlab = 'Age of Facebook users',
      ylab = 'Number of users in sample',
      color = I('black'), fill = I("#FF9999")) +
    scale_x_continuous(lim=c(13,80), breaks = seq(13,80,10))
```

## Transforming Data
## Scaled data

```{r}
library(gridExtra)
p1 = qplot(x = friend_count, data = pf)
p2 = qplot(x = log10(friend_count +1), data = pf)
p3 = qplot(x = sqrt(friend_count), data = pf)
# arrange plots in grid
grid.arrange(p1, p2, p3, ncol = 3)
```

## Frequency Polygons (before we had histograms)
```{r}
qplot(x = friend_count, data = subset(pf, !is.na(gender)), biwidth = 10) +
    scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 50)) +
    facet_wrap(~gender)

qplot(x = friend_count, y = ..count../sum(..count..),
      data = subset(pf, !is.na(gender)),
      xlab = "Friend Count",
      ylab = "Proportion of Users with that friend count",
      biwidth = 10, geom = "freqpoly", color = gender) +
    scale_x_continuous(lim = c(0, 1000), breaks = seq(0, 1000, 50))
```

## Frequency Polygons (who has more likes)
```{r}
qplot(x = www_likes, 
      data = subset(pf, !is.na(gender)),
      geom = "freqpoly", color = gender) +
    scale_x_continuous() +
    scale_x_log10()
```


## Statistics of likes 'by' Gender

```{r}
by(pf$www_likes, pf$gender, sum)
```

## Box Plots
```{r}
qplot(x = gender, y = friend_count,
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot') +
    coord_cartesian(ylim = c(0,250))

by(pf$friend_count, pf$gender, summary)
```

## Box Plots, Quartiles, Friend Requests
```{r}
qplot(x = gender, y = friendships_initiated,
      data = subset(pf, !is.na(gender)),
      geom = 'boxplot') +
    coord_cartesian(ylim = c(0,130))

by(pf$friendships_initiated, pf$gender, summary)
```

## Getting logical
```{r}
summary(pf$mobile_likes)

summary(pf$mobile_likes > 0)

mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0,1,0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
summary(pf$mobile_check_in)
```

## Price Histogram
```{r}
library(ggplot2)
qplot(price, data = diamonds,
      xlab = 'Price of diamond',
      ylab = 'Number of diamonds in sample',
      color = I('red'), fill = I("#FF9999"))

summary(diamonds$price)

```

## Cheaper diamonds
```{r}
library(ggplot2)
qplot(price, data = diamonds,
      xlab = 'Price of diamond',
      ylab = 'Number of diamonds in sample',
      color = I('red'), fill = I("#FF9999"),
      binwidth = 1) +
  scale_x_continuous(lim = c(0, 3000), breaks = seq(0, 3000, 250))

ggsave('priceHistogram.png')
#There are no diamonds that cost $1500.
#For diamonds that cost less than $2,000, the most common price of a diamond is around $700 with the mode being $605 (binwidth = 1). 
```

## Price by Cut Histogram
```{r}
library(ggplot2)
a<-qplot(price, data = diamonds,
      xlab = 'Price of diamond',
      ylab = 'Number of diamonds in sample',
      color = I('red'), fill = I("#FF9999")) +
      facet_wrap(~cut, ncol = 2)
by(diamonds$price,diamonds$cut,max)
by(diamonds$price,diamonds$cut,min)
by(diamonds$price,diamonds$cut,median)
```

## Scales and Multiple Histograms
```{r}
qplot(x = price, data = diamonds,
      xlab = 'Price of diamond',
      ylab = 'Number of diamonds in sample',
      color = I('red'), fill = I("#FF9999")) +
      facet_wrap(~cut, scales = "free_y")
by(diamonds$price, diamonds$cut, summary)
```

## Price per Carat by Cut
```{r}
qplot(x = price/carat, data = diamonds,
      xlab = 'Price per carat',
      ylab = 'Number of diamonds in sample',
      color = I('red'), fill = I("#FF9999"),
      binwidth = 50) +
      facet_wrap(~cut)


qplot(x = price/carat, data = diamonds,
      xlab = 'Price per carat',
      ylab = 'Number of diamonds in sample',
      color = I('red'), fill = I("#FF9999"),
      binwidth = 1) +
      facet_wrap(~cut) +
      scale_x_log10()
```


## Price Box Plots by clarity
```{r}
qplot(x = clarity, y = price,
      data = subset(diamonds, !is.na(clarity)),
      geom = 'boxplot') 

by(diamonds$price, diamonds$clarity, summary)
ggsave('priceHistogrambyclarity.png')
```

## Price Box Plots by color
```{r}
qplot(x = color, y = price,
      data = subset(diamonds, !is.na(color)),
      geom = 'boxplot') 

by(diamonds$price, diamonds$color, summary)
IQR(subset(diamonds, color == "D")$price)
IQR(subset(diamonds, color == "J")$price)
ggsave('priceHistogrambycolor.png')
```

## Carat Frequency Polygon
```{r}
qplot(carat, data = diamonds, geom = "freqpoly", binwidth = 0.01)
   
```



