Lesson 5
========================================================

# Third Qualitative Variable

```{r}
setwd("D:/R/Udacity/EDA_Course_Materials/lesson5")
```

```{r}
pf <- read.csv('pseudo_facebook.tsv', sep = '\t')
library(ggplot2)
library(dplyr)

ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() +
    stat_summary(fun.y = mean, geom = 'point', shape = 4)
```

```{r}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) + 
    geom_line(aes(color = gender), stat = 'summary', fun.y = median)
```

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:
```{r}
pf.fc_by_age_gender <- subset(pf, !is.na(gender)) %>%
    group_by(age, gender) %>%
    summarise(mean_friend_count = mean(friend_count), median_friend_count = median(friend_count), n = n()) %>%
    ungroup() %>%
    arrange(age)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = median_friend_count),
       data = pf.fc_by_age_gender) + 
    geom_line(aes(color = gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
library(reshape2)
pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender, age ~ gender, value.var = 'median_friend_count')
head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}
ggplot(aes(x = age, y = female / male),
data = pf.fc_by_age_gender.wide) +
    geom_line() +
    geom_hline(yintercept = 1, alpha = 0.3, linetype = 2)
```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
#        (2004, 2009]
#        (2009, 2011]
#        (2011, 2012]
#        (2012, 2014]

summary(pf$year_joined)
table(pf$year_joined)

pf$year_joined.bucket <- cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014))
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
table(pf$year_joined.bucket, useNA = 'ifany')

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) + 
    geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(year_joined.bucket))) + 
    geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) +
    geom_line(stat = 'summary', fun.y = mean, linetype = 2)
```

***

### Friending Rate
Notes:

```{r Friending Rate}
with(subset(pf, tenure >= 1), summary(friend_count / tenure))
```

***

### Friendships Initiated
Notes:

What is the median friend rate?
0.2205

What is the maximum friend rate?
417.0000

```{r Friendships Initiated}
ggplot(aes(x = tenure, y = friendships_initiated / tenure), data = subset(pf, tenure >=1)) +
    geom_line(aes(color = year_joined.bucket),)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)


ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_smooth(aes(color = year_joined.bucket))
```

### Introducing the Yogurt Data Set
Notes:
Yogurt data has many rows per household, one for each purchase occasion.
***

### Histograms Revisited
Notes:

There are some observation which almost don't have purchases.

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
str(yo)

# Change the id from an int to a factor
yo$id <- factor(yo$id)
str(yo)

qplot(data = yo, x = price, fill = I('#F79420'))
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}
summary(yo)
length(unique(yo$price))
table(yo$price)

yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)

summary(yo$all.purchases)
```

***

### Prices over Time
Notes:

Model of most common price encreases over time.

```{r Prices over Time}
qplot(x = all.purchases, data = yo, binwidth = 1, fill = I('#099DD9'))

ggplot(aes(x = time, y = price), data = yo) +
    geom_jitter(alpha = 1/4, shape = 21, fill = I('#F79420'))
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(666)
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x = time, y = price),
       data = subset(yo, id %in% sample.ids)) +
    facet_wrap(~ id) +
    geom_line() +
    geom_point(aes(size = all.purchases), pch = 18, color = I('#7D26CD'))
ggsave('HH_sample.jpeg', width = 10, height = 10)
```

***

## Scatterplot Matrices
```{r}
library(GGally)
theme_set(theme_minimal(20)) 

set.seed(1836)
pf_subset <- pf[, c(2:15)]
names(pf_subset)
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ])
```


### Scatterplot Matrix
Notes:
Scatterplots are below the diagonal, and categorical variables, like gender, create faceted histograms.
***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

