Unemployment in the World
====================================================================

```{r}
setwd("D:/R/Udacity/EDA_Course_Materials/lesson5")
library(ggplot2)
library(tidyr)
library(dplyr)
```

```{r}
# read 6 files

# female
f_15_24 <- read.csv("indicator_f 15-24 unemploy - Data.csv", header = TRUE, row.names = 1)
f_25_54 <- read.csv("indicator_f 25-54 unemploy - Data.csv", header = TRUE, row.names = 1)
f_above_55 <- read.csv("indicator_f above 55 unemploy - Data.csv", header = TRUE, row.names = 1)

# male
m_15_24 <- read.csv("indicator_m 15-24 unemploy - Data.csv", header = TRUE, row.names = 1)
m_25_54 <- read.csv("indicator_m 25-54 unemploy - Data.csv", header = TRUE, row.names = 1)
m_above_55 <- read.csv("indicator_m above 55 unemploy - Data.csv", header = TRUE, row.names = 1)
```

```{r}
#add new columns (sex,age,country) to datasets
f_15_24['sex'] = 'female'
f_25_54['sex'] = 'female'
f_above_55['sex'] = 'female'

f_15_24['age'] = '15-24'
f_25_54['age'] = '25-54'
f_above_55['age'] = 'above 55'

f_15_24['country'] = row.names(f_15_24)
f_25_54['country'] = row.names(f_25_54)
f_above_55['country'] = row.names(f_above_55)

m_15_24['sex'] = 'male'
m_25_54['sex'] = 'male'
m_above_55['sex'] = 'male'

m_15_24['age'] = '15-24'
m_25_54['age'] = '25-54'
m_above_55['age'] = 'above 55'

m_15_24['country'] = row.names(m_15_24)
m_25_54['country'] = row.names(m_25_54)
m_above_55['country'] = row.names(m_above_55)
```

```{r}
# merge datasets
data <- rbind(f_15_24, f_25_54, f_above_55, m_15_24, m_25_54, m_above_55)
```

```{r}
#making tidydata 
t_data <- gather(data, year, total, -sex, -age, -country)
```

```{r}
#correct year names
t_data['year'] <- extract_numeric(t_data$year)
```

## First look at tidy_data
```{r}
head(t_data)
dim(t_data)
summary(t_data)
```

## Unemployment rate according to gender
Notes:
Unemployment quantiles and means for male and female are alomost the same but female boxplot has much higher outliers even over 50% of unemployment.
```{r}
ggplot(aes(x = sex, y = total),
       data = subset(t_data, !is.na(total))) + geom_boxplot() +
    stat_summary(fun.y = mean, geom = 'point', shape = 2)
```


## Unemployment *mean* rate according to gender
Notes:
Mean graphs for 1980 - 2005 show that in average female have higher or equal to male unemployment rate.
```{r}
ggplot(aes(x = year, y = total),
       data = subset(t_data, !is.na(total))) + 
    geom_line(aes(color = sex), stat = 'summary', fun.y = mean)
```
 
## Unemployment rate of each country yearly by age
Notes:
Check how different are patterns for different countries. 
We can aslo see that Spain is the most problematic country in our dataset.
```{r}
t_data$age <- factor(t_data$age)
t_data$age <- ordered(t_data$age, levels = c('15-24','25-54', 'above 55'))

ggplot(t_data, aes(x = year, y = total)) +
    geom_point(aes(color = t_data$age), alpha = (1/2), position = position_jitter(width = 0.4)) +
    facet_wrap( ~ country) +
    scale_color_brewer(type = 'qual')
```

## Unemployment analysis by age 
Notes:
Unemployment is higher in younger age.
```{r}
ggplot(t_data, aes(x = year, y = total, color = age)) +
    geom_point() +
    scale_color_brewer(type = 'qual')
```
